Human capital mediates natural selection in contemporary humans

### 2022-05-30 why-natural-selection.Rmd
Use condensed font in graphs

knitr::opts_chunk$set(
        echo       = FALSE, 
        warning    = FALSE, 
        cache      = FALSE, 
        error      = FALSE,
        fig.height = 3.5
      )

@@

theme_set(theme_minimal())

@@

pgs_over_time %>% 
      mutate(score_name = pretty_names(score_name)) %>% 
      ggplot(aes(YOB, group = score_name, colour = Change, shape = Change)) + 
      geom_linerange(aes(ymin = score_lo, ymax = score_hi), color = "grey70", 
                       size = 3) +
      geom_point(aes(y = score), size = 1.5) +
      facet_wrap(~score_name, ncol = 5) +
      my_hline +
      scale_color_manual(values = c("-" = "darkred", "+" = "darkgreen", "o" = "black")) +
      scale_shape_manual(values = c(
                                      "-" = "triangle down open", 
                                      "+" = "triangle open", 
                                      "o" = "circle open"
                                    )) +
      theme(
        line        = element_line(size = 0.5, lineend = "round"),
        strip.text  = element_text(size = 6), 
        axis.text.x = element_text(size = 7),
        legend.position = "none"
      ) +
      labs(y = "PGS", x = "Birth year (5 year intervals)")

@@

dfr %>% 
      mutate(
        PGS      = ifelse(abs(effect_size) > 0.027, PGS, ""),
        PGS      = sub("(^\\w*?)_.*", "\\1", PGS)
      ) %>% 
      filter(
        Controls == "None"
      ) %>% 
      ggplot(aes(effect_size, cor)) + 
        geom_point(color = "steelblue4") +
        geom_text(aes(label = PGS), size = 3, colour = "black", 
                    check_overlap = TRUE, nudge_x = 0.008) +
        facet_wrap( ~ Correlation, scales = "free") + 
        my_vline + 
        my_hline +
        labs(x = "Selection effect", y = "Correlation") +
        coord_cartesian(clip = "off") +
        theme(
          legend.position = "none",
          panel.spacing = unit(20, "pt")
        )

@@

res_mediation %>% 
      mutate(
        term = pretty_names(term),
        term = fct_reorder(term, prop_ind, order_abs(1)),
        `Indirect effect p < 0.05/23` = sig 
      ) %>% 
      ggplot(aes(y = term)) + 
        geom_col(aes(x = prop_ind, fill = `Indirect effect p < 0.05/23`), 
                   width = .5) + 
        geom_errorbar(aes(
                          xmin = prop_ind_conf_low, 
                          xmax = prop_ind_conf_high,
                          linetype = "Bootstrap 95% c.i. uncorrected",
                        ),
                        width = 0.2,
                        color = "grey50"
                      ) +
        scale_fill_manual(values = c("FALSE" = "grey65", "TRUE" = "darkgreen")) +
        my_vline +
        scale_x_continuous(labels = scales::percent, n.breaks = 7) +
        coord_cartesian(xlim = c(-0.5, 1)) +
        labs(x = "Proportion mediated", y = "") +
        scale_linetype_manual(name = "", values = c("Bootstrap 95% c.i. uncorrected" = 1)) +
        my_theme(
          panel.grid.major.y = element_blank(),
          panel.grid.minor.x = element_blank()
        )

@@

res_pcs %>% 
      mutate(
        estimate = abs(estimate),
        dep.var  = fct_relevel(dep.var, "RLRS"),
        dep.var  = fct_recode(dep.var,
          "Parents"     = "RLRS_parents",
          "Respondents" = "RLRS"
        ),
        top_pcs  = ifelse(as.numeric(pc_name) <= 5, term, "")
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(estimate, dep.var, shape = p < 0.05/{{n_regs}})) + 
        geom_point(alpha = 0.8, color = "steelblue4", position = pj) +
        geom_text(aes(estimate, dep.var, label = top_pcs), position = pj, 
                    hjust = "right", color = "black", size = 3) +
        scale_x_log10(labels = scales::label_number(accuracy = 0.0001)) +
        labs(x = "Effect size (log scale)", y = "") + 
        scale_shape_manual(values = c("FALSE" = "circle open", "TRUE" = "circle"))

>>>>

knitr::opts_chunk$set(
        echo       = FALSE, 
        warning    = FALSE, 
        cache      = FALSE, 
        error      = FALSE,
        fig.height = 3.5,
        dev        = "cairo_pdf"
      )

@@

theme_set(theme_minimal())
theme_update(
  text = element_text(family = "Abadi MT Condensed Light")
)

@@

pgs_over_time %>% 
      mutate(score_name = pretty_names(score_name)) %>% 
      ggplot(aes(YOB, group = score_name, colour = Change, shape = Change)) + 
      geom_linerange(aes(ymin = score_lo, ymax = score_hi), color = "grey70", 
                       size = 3) +
      geom_point(aes(y = score), size = 1.5) +
      facet_wrap(~score_name, ncol = 5) +
      my_hline +
      scale_color_manual(values = c("-" = "darkred", "+" = "darkgreen", "o" = "black")) +
      scale_shape_manual(values = c(
                                      "-" = "triangle down open", 
                                      "+" = "triangle open", 
                                      "o" = "circle open"
                                    )) +
      theme(
        line        = element_line(size = 0.5, lineend = "round"),
        strip.text  = element_text(size = 8), 
        axis.text.x = element_text(size = 7),
        legend.position = "none"
      ) +
      labs(y = "PGS", x = "Birth year (5 year intervals)")

@@

dfr %>% 
      mutate(
        PGS      = ifelse(abs(effect_size) > 0.027, PGS, ""),
        PGS      = sub("(^\\w*?)_.*", "\\1", PGS)
      ) %>% 
      filter(
        Controls == "None"
      ) %>% 
      ggplot(aes(effect_size, cor)) + 
        geom_point(color = "steelblue4") +
        geom_text(aes(label = PGS), size = 3, colour = "black", 
                    family = "Abadi MT Condensed Light",
                    check_overlap = TRUE, nudge_x = 0.004) +
        facet_wrap( ~ Correlation, scales = "free") + 
        my_vline + 
        my_hline +
        labs(x = "Selection effect", y = "Correlation") +
        coord_cartesian(clip = "off") +
        theme(
          legend.position = "none",
          panel.spacing = unit(20, "pt")
        )

@@

res_mediation %>% 
      mutate(
        term = pretty_names(term),
        term = fct_reorder(term, prop_ind, order_abs(1)),
        `Indirect effect p < 0.05/23` = sig 
      ) %>% 
      ggplot(aes(y = term)) + 
        geom_col(aes(x = prop_ind, fill = `Indirect effect p < 0.05/23`), 
                   width = .5) + 
        geom_errorbar(aes(
                          xmin = prop_ind_conf_low, 
                          xmax = prop_ind_conf_high,
                          linetype = "Bootstrap 95% c.i. uncorrected",
                        ),
                        width = 0.2,
                        color = "grey50"
                      ) +
        scale_fill_manual(values = c("FALSE" = "grey65", "TRUE" = "darkgreen")) +
        my_vline +
        scale_x_continuous(labels = scales::percent, n.breaks = 7) +
        coord_cartesian(xlim = c(-0.5, 1)) +
        labs(x = "Proportion mediated", y = "") +
        scale_linetype_manual(name = "", 
          values = c("Bootstrap 95% c.i. uncorrected" = 1)) +
        my_theme(
          panel.grid.major.y = element_blank(),
          panel.grid.minor.x = element_blank(),
          legend.justification = "top"
        )

@@

res_pcs %>% 
      mutate(
        estimate = abs(estimate),
        dep.var  = fct_relevel(dep.var, "RLRS"),
        dep.var  = fct_recode(dep.var,
          "Parents"     = "RLRS_parents",
          "Respondents" = "RLRS"
        ),
        top_pcs  = ifelse(as.numeric(pc_name) <= 5, term, "")
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(estimate, dep.var, shape = p < 0.05/{{n_regs}})) + 
        geom_point(alpha = 0.8, color = "steelblue4", position = pj) +
        geom_text(aes(estimate, dep.var, label = top_pcs), position = pj, 
                    hjust = "right", color = "black", size = 3,
                    family = "Abadi MT Condensed Light") +
        scale_x_log10(labels = scales::label_number(accuracy = 0.0001)) +
        labs(x = "Effect size (log scale)", y = "") + 
        scale_shape_manual(values = c("FALSE" = "circle open", "TRUE" = "circle"))
